---
title: "Untitled"
author: "Taryn McLaughlin"
date: "5/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(GGally)
library(gplots)
library(ggfortify)
library(RColorBrewer)
library(data.table)
library(ggpubr)
library(pcaMethods)
library(cluster)
library(knitr)
library(heatmap.plus)
library(stats)
#Generate some data so you don't have to use my actual data because it's a mess
data(diamonds)
dfd<-diamonds
dfd<-dplyr::select(diamonds, -cut, -color, -clarity)
rows<-sample(1:53940, 100)
dfd<-dfd[rows,]
dfd<-data.matrix(dfd)

lum<-read.csv("/Users/tarynam/Desktop/TH1 TH2_Taryn/Luminex_combined_background.txt")
lum<-dplyr::select(lum, -X)
lum$order<-paste(lum$SM, lum$Stim) 
lum<-dplyr::arrange(lum, order) #so that the matrix is ordered first by SM and then within by Stim
rownames(lum)<-lum$Sample
```

##Hierarchical Clustering
Uses defaults of complete clustering using euclidean distance
```{r}
df<-data.matrix(lum[,2:10])
hh <- dist(df) %>% hclust
plot(as.dendrogram((hh)))


dataMatrixOrdered <- df[hh$order, ]
par(mfrow = c(1, 3))

## Complete data
image(t(dataMatrixOrdered)[, nrow(dataMatrixOrdered):1])

## Show the row means
plot(rowMeans(dataMatrixOrdered), 216:1, xlab = "Row Mean", ylab = "Row", pch = 19)
## Show the column means
plot(colMeans(dataMatrixOrdered), xlab = "Column", ylab = "Column Mean", pch = 19)

```

##K-means Clustering
```{r}
set.seed(1234)
df<-lum[,2:10]
km <- kmeans(df, 10, 15)
lum$km<-factor(km$cluster)
ggplot(lum, aes(x=Stim, y=km, col=SM))+geom_jitter(width=.3, height = 0, aes(alpha=0.5), size=2)

par(mfrow = c(1, 2))
image(t(df)[, nrow(df):1], yaxt = "n", main = "Original Data")
image(t(df)[, order(km$cluster)], yaxt = "n", main = "Clustered Data")
```

##Principal Components Analysis
```{r fig.width=12}
df<-lum[,2:10]
pca1 <- prcomp(df, scale = TRUE)
pcas<-data.frame(pca1$x)
rotations<-data.frame(pca1$rotation)
rotations$cytokine<-row.names(rotations)
melt<-melt(rotations, id.vars="cytokine")

stims<-factor(lum$Stim)
worms<-factor(lum$SM)
ggplot(pcas, aes(x=pcas[,1], y=pcas[,2]))+geom_point(aes(col=worms))+labs(y="PCA2", x="PCA1")
kable(rotations[,1:9])
ggplot(melt, aes(x=cytokine, y=value, fill=variable))+geom_col()
ggplot(melt, aes(x=variable, y=value, fill=cytokine))+geom_col()

pca2 <- pca(df, scale = "uv", center = T, nPcs = 2, method = "svd")
slplot(pca2, scoresLoadings = c(T,T))

autoplot(prcomp(df), data = lum, colour = 'Stim',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3)
```

#Common Factor Analysis
setting the number of factors is completely arbitrary so you want to try a few and then look at the p-value for each. This p-value is for the hypothesis that the factor split model built is a "perfect fit". So lower means it doesn't fit well. There are caps on how many factors you can use based on the number of variables.

```{r}
nums<-1:5
pvals<-c(factanal(df, 1)$PVAL, factanal(df, 2)$PVAL, factanal(df, 3)$PVAL, factanal(df, 4)$PVAL, factanal(df, 5)$PVAL)
kable(cbind(nums, pvals))
```

We pick the 5-level one and explore the output.

```{r}
five<-factanal(df, 5)
print(five, digits = 2, cutoff = .2, sort = TRUE)
```
